---
title: '`Week 2` Exploratory Data Analysis'
author: '`r if(knitr::is_html_output()) {"&#x1f468;&#x1F3FB;&#x200d;&#x1f4bb; Anderson H Uyekita"} else {NULL}`'
output:
  github_document: default
params:
  author: "Anderson H Uyekita"
  course: "Exploratory Data Analysis"
  course_url: "https://www.coursera.org/learn/exploratory-data-analysis"
  specialization: "Data Science: Foundations using R Specialization"
  specialization_url: "https://www.coursera.org/specializations/data-science-foundations-r"
  instructor: "Roger D Peng"
  week: "Week 2"
  course_start: !r lubridate::ymd("2022/06/07")
  course_finish: !r lubridate::today()
  slug: 'Week%202'
---

```{r setup, include=FALSE,message=FALSE,warning=FALSE,cache=TRUE}
# Defining document chunks options
knitr::opts_chunk$set(echo = TRUE,cache=TRUE,message=FALSE,warning=FALSE)

# Loading required libraries
library(rmarkdown)
library(lubridate)
```

* `r if(!knitr::is_html_output()) {sprintf(fmt = "&#x1f468;&#x1F3FB;&#x200d;&#x1f4bb; Author: %s", params$author)}`
* `r sprintf(fmt = "&#x1f4da; Specialization: [%s](%s){target='_blank' rel='noopener'}", params$specialization, params$specialization_url)`
* `r sprintf(fmt = "&#x1f4d6; Course: [%s](%s){target='_blank' rel='noopener'}", params$course, params$course_url)`
    * `r sprintf(fmt = "&#x1F9D1;&#x200d;&#x1F3EB; Instructor: %s", params$instructor)`
* `r sprintf(fmt = "&#x1F4C6; %s", params$week)`
    * `r sprintf(fmt = "&#x1F6A6; Start: %s", format(params$course_start, "%A, %d %B %Y"))`
    * `r sprintf(fmt = "&#x1F3C1; Finish: %s", format(params$course_finish, "%A, %d %B %Y"))`

--------------------------------------------------------------------------------

#### Assignments & Deliverables

* &#x1F4BB; Swirl
    * Lattice Plotting System
    * Working with Colors
    * GGPlot2 Part1
    * GGPlot2 Part2
    * GGPlot2 Extras
* [&#x1F4DD; Quiz 1](./quiz-2_exploratory-data-analysis.md)

--------------------------------------------------------------------------------

#### Slides

* Lesson 1: Lattice Plotting System [](){id="lesson-1"}
    * [Lattice Plotting System](./slides/2_1_plotting-lattice.pdf)
* Lesson 2: ggplot2 [](){id="lesson-2"}
    * [ggplot2](./slides/2_2_ggplot-2.pdf)

--------------------------------------------------------------------------------

## Class Notes

### [<kbd>Lesson 1</kbd>](#lesson-1) Lattice Plotting System in R

**The Lattice Plotting System**

The lattice plotting system is implemented using `lattice` and `grid` packages.

> * lattice: contains code for producing Trellis graphics, which are independent of the “base” graphics system; includes functions like `xyplot`, `bwplot`, `levelplot`
> * grid: implements a different graphing system independent of the “base” system; the lattice package builds on top of grid
> * The lattice plotting system does not have a "two-phase" aspect with separate plotting and annotation like in base plotting
> * All plotting/annotation is done at once with a single function call

**Lattice Functions**

> * `xyplot`: this is the main function for creating scatterplots
> * `bwplot`: box-and-whiskers plots (“boxplots”)
> * `histogram`: histograms
> * `stripplot`: like a boxplot but with actual points
> * `dotplot`: plot dots on "violin strings"
> * `splom`: scatterplot matrix; like `pairs` in base plotting system
> * `levelplot`, contourplot: for plotting "image" data

Lattice functions generally take a formula for their first argument, usually of the form

```
xyplot(y ~ x | f * g, data)
```
We use the formula notation here, hence the `~` on the left of the `~` is the y-axis variable, on the right is the x-axis variable.

```{r}
library(datasets)
library(lattice)

# Scatter plot example.
xyplot(x = mpg ~ wt, data = mtcars)
```

`f` and `g` are conditioning variables — they are optional, the `*` indicates an interaction between two variables.

```{r}
# Scatter plot example with facets.
xyplot(x = mpg ~ wt | factor(cyl), data = mtcars)
```

The second argument is the data frame or list from which the variables in the formula should be
looked up.

* If no data frame or list is passed, then the parent frame is used.
* If no other arguments are passed, there are defaults that can be used.

**Lattice Behavior**

Lattice functions behave differently from base graphics functions in one critical way.

> * Base graphics functions plot data directly to the graphics device (screen, PDF file, etc.)
> * Lattice graphics functions return an object of class trellis
> * The print methods for lattice functions actually do the work of plotting the data on the graphics device.
> * Lattice functions return "plot objects" that can, in principle, be stored (but it’s usually better to just save the code + data).
> * On the command line, trellis objects are *auto-printed* so that it appears the function is plotting the data

**Summary**

> * Lattice plots are constructed with a single function call to a core lattice function (e.g. xyplot)
> * Aspects like margins and spacing are automatically handled and defaults are usually sufficient
> * The lattice system is ideal for creating conditioning plots where you examine the same kind of plot under many different conditions
> * Panel functions can be specified/customized to modify what is plotted in each of the plot panels


### [<kbd>Lesson 2</kbd>](#lesson-2) Plotting with ggplot2


**Summary**
